Invasive lionfish present region-wide variation of allometric growth in the

### 2018-05-29 docs/Manuscript.Rmd
Incorporated Sean's comments

---
title: Invasive lionfish present region-wide variation of allometric growth in the
  Western Atlantic
author: "Juan Carlos Villaseñor-Derbez$^1$, Sean Fitzgerald"
header-includes:
- \usepackage{float}
- \floatplacement{figure}{H}
output:
  pdf_document:
    keep_tex: yes
    number_sections: no
    template: peerj_template.tex
institution: "Bren School of Environmental Sciences and Management, University of California Santa Barbara, Santa Barbara, California, U.S."
abstract: "Lionfish (*Pterois volitans/miles*) are an invasive species in the Western Atlantic and the Caribbean. In order to better manage the invasion, we must be able to accurately estimate their total biomass. This work compares length-weight relationships of the invasive lionfish through the invasion range. A review of 17 length-weight relationships reported in 12 peer-reviewed studies shows that lionfish exhibit spatial variation in weight-at-length. The reviewed parameters indicate that, for the same length, lionfish in the Caribbean have lower body mass than in the Atlantic or Gulf of Mexico. This highlights the importance of using site-specific parameters to estimate biomass from length observations. This study also reports a new pair of length-weight parameters ($a = 3.2056\times 10^{-6}; b = 3.235$) for organisms sampled in the central Mexican Caribbean region. These findings can have major implications in management, especially when estimating biomass available for harvest, predicting effects on local ecosystems, or evaluating the effectiveness of removal programs."
keywords: "A, B, C"
---

@@

At least 84% of marine eco-regions have reported the presence of an invasive species \citep{molnar_2008}, which represent a major threat to local biodiversity and the economic activities that depend on it \citep{bax_2003}. Invasive species may threaten native species through predation, competition, or indirect habitat effects \citep{davis_2003, gurevitch_2004}. By 2005, the economic cost of invasive species to the United States was  estimated at USD \$120 billion per year \citep{pimentel_2005}.

Lionfish (*Pterois volitans/miles* complex) are an invasive species in the western Atlantic and the Caribbean, likely introduced through liberation of aquarium-kept organisms \citep{betancurr_2011}. They are the first invasive marine vertebrates established along the North Atlantic Caribbean coasts \citep{schofield_2009,schofield_2010,sabidoitza_2016}. Lionfish have been widely reported in coral reefs \citep{aguilarperera_2010}, but also in other habitats such as estuaries \citep{jud_2011}, mangroves \citep{barbour_2010}, hard-bottomed areas \citep{muoz_2011}, and mesophotic reefs \citep{andradibrown_2017}. Their presence in these waters has been labeled as a major marine invasion because they threaten local biodiversity, spread rapidly, and are difficult to manage \citep{hixon_2016}.

A substantial amount of research has been done to describe lionfish feeding ecology in North Carolina \citep{muoz_2011}, the Bahamas \citep{morris_2009,cote_2013}, Northern Gulf of Mexico \citep{dahl_2014}, Mexican Caribbean \citep{valdezmoreno_2012,villaseorderbez_2014}, Belize \citep{hackerott_2017}, and Costa Rica \citep{sandel_2015}. \citet{peake_2018} show that invasive lionfish prey on at least 167 different species across the tropical and temperate North Atlantic. Their feeding behavior and high consumption rates can reduce recruitment and population sizes of native reef-fish species, and can further endanger reef fish \citep{albins_2008, green_2012,rocha_2015}. (However, see \citet{hackerott_2017} for a case where there was no evidence that lionfish affected the density, richness, or community composition of prey fishes). Major efforts have been made to understand the possible impacts of the invasion by keeping track of its range through time \citep{schofield_2009,schofield_2010} and predicting invasion ranges under future climates \citep{grieve_2016}. By combining information from these disciplines, researchers have been able to predict the trophic impacts of lionfish \citep{ariasgonzalez_2011}, which can then be translated into ecosystem-level and economic impacts.

Governments and non-profit organizations have sought to reduce lionfish densities through removal programs and incentivizing its consumption \citep{chin_2016}. In some cases, these have shown to significantly reduce --but not quite eliminate-- lionfish abundances at local scales \citep{sandel_2015,chin_2016,deleon_2013}. In addition, culling programs can help stabilize or grow native prey fish populations \citep{cote_2014}. Complete eradication of lionfish through fishing is unlikely because of their rapid recovery rates and ongoing recruitment to shallow-water areas from their persistent populations in mesophotic coral ecosystems \citep{barbour_2011,andradibrown_2017}. However, promoting its consumption might create a level of demand capable of sustaining a stable fishery, which can help control shallow-water populations while providing alternative livelihoods and avoiding further impacts to local reef biota \citep{chin_2016}.

The feasibility of establishing fisheries through lionfish removal programs has been extensively evaluated through field observations and empirical modeling \citep{barbour_2011,morris_2011,deleon_2013,johnston_2015,sandel_2015,chin_2016,usseglio_2017}. One contributing factor to the success of many removal programs is the sedentary nature of adult lionfish \citep{jud_2012}. Culling programs are effective in reducing adult populations largely because lionfish exhibit high levels of site fidelity and rarely leave their home range in most cases \citep{Fishelson_1997,cote_2014,kochzius_2005}. As a result of this sedentary behavior, lionfish are also likely to exhibit high levels of spatial variation in important life history characterstics such as growth or natural mortality rate \citep{hutchinson_2008,wilson_2012}. The importance of considering spatial heterogeneity is well-documented in terms of assessing and managing sedentary species \citep{gunderson_2008,guan_2013}, and such variation should be accounted for when evaluating the feasibility of establishing lionfish fisheries as well.

Empirical modeling efforts examining the feasibility of establishing fisheries for lionfish involve modeling changes in biomass in response to changes in mortality (*i.e.* culling). A common way to model this is via age- or length-structured population models \citep{cote_2014,barbour_2011,andradibrown_2017}, which convert fish length or age to weight to then calculate total biomass. Therefore, length-weight relationships are an essential component of these models. However, the length-weight relationship can vary across regions as a response to biotic (*e.g.* local food availability) and abiotic (*e.g.* water temperature) conditions \citep{johnson_2016}. Literature suggests that site-specific parameters are necessary to obtain accurate biomass estimates when length-weight relationships present spatial variation. This becomes ingreasingly important when estimating the potential effectiveness (and resources neded) of lionfish culling programs or identifying total biomass available for harvest by fishers \citep{barbour_2011,morris_2011,johnston_2015,chin_2016,cote_2014}. Genetic analysis of invasive lionfish suggest that two distinct population groups exist \citep{betancurr_2011}, and as such we would expect their biology to differ. Despite the large number of studies reporting site-specific length-weight relationships, no studies have described region-wide differences in these.

The objective of this paper is to describe the spatial pattern of length-weight relationships of lionfish in the Caribbean and Western Atlantic and identify wether these differences are trivial. Length-weight relationships for lionfish exist for North Carolina, Northern and Southern Gulf of Mexico, the Southern Mexican Caribbean, Bahamas, Little Cayman, Jamaica, Bonaire, Puerto Rico, and Costa Rica, but remain unavailable for the central Mexican Caribbean \citep{barbour_2011,fogg_2013,dahl_2014,aguilarperera_2016,sabidoitza_2016,sabidoitz_2016,darling_2011,edwards_2014,chin_2016,deleon_2013,toledohernndez_2014,sandel_2015}. This study also provides the first length-weight relationship for this region.

@@

The main objective of this work was to compare allometric growth of lionfish throughout their invasion range. Allometric parameters were retreived from scientific literature, and an additional pair of parameters were calculated from filed observations in the central Mexican Caribbean.

Length-weight relationships (n = `r dim(review)[1]-1`) identified in literature were obtained for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and Caribbean (n = 10). Information on sampling methods, sex differentiation, location, and depth ranges of each study was retrieved when available. Studies were assumed to include both genders if gender was unspecified. Studies inconsistently defined $a$ as either the ponderal index from Eq. \ref{eq:allometric} or the y-intercept ($c$) from Eq. \ref{eq:log-alo-trans}. Other studies incorrectly reported parameters as mm-to-g conversions when they were in fact cm-to-g conversions. Here, all parameters are reported as TL(mm) to TW(gr) conversions. When required, values from other studies were transformed for consistency. Locations where allometric studies have been performed are shown in Figure \ref{fig:map} and Table \ref{tab:all_params}.

@@

Parameters from the central Mexican Caribbean were obtained from data collected in `r length(unique(fish$Location))` sampling sites along the central Mexican Caribbean coast in 2010 (Table S1). Sampling locations included wall and carpet reefs at depths between `r min(fish$Depth_m)` m and `r max(fish$Depth_m)` m.  All observed lionfish (n = `r dim(fish)[1]`) were collected using hand nets and numbered collection bottles. The use of hand nets prevented any weight loss due to bleeding and allowed better representation of small sizes by eliminating gear selectivity. Organisms were euthanized via pithing and Total Length (TL; mm) and Total Weight (TW; g) were recorded before freezing organisms.

@@

\begin{equation}
\label{eq:log-alo-trans}
Y = mX + c
\end{equation}

Where $Y = log_{10}(TW)$, $m = b$, $X = log_{10}(TL)$, and $c = log_{10}(a)$. Since $b = m$, we will only use $b$ throughout the paper for simplicity. The coefficients ($c$ and $b$) were estimated with  an Ordinary Least Squares Regression and heteroskedastic-robust standard error correction \citep{zeileis_2004}. The $b$ coefficient was tested against the null hypothesis of isometric growth (*i.e.* $H_0: b = 3$). Coefficients were tested with a two-tailed Student's t, and the significance of the regression was corroborated with an F-test.

@@

Combining the length-weight parameters extracted from the literature and the additional pair calculated here, we obtain a total of `r dim(review)[1]` pairs. We test wether using *ex situ* parameters produces differences in weight estimates by estimating TW for the 109 length observations of lionfish from the central Mexican Caribbean with each of the `r dim(review)[1]` relationships. Predicted weights were divided by know observed weights to obtain a simple measurement of over- or underestimation. Difference in mean weight ratios across studies were tested with a one-way analysis of variance (ANOVA). All analyses were performed in R version `r paste(version$major, version$minor, sep = ".")` \citep{rcore_2018}. Raw data and code used in this work are available at dryad.org.

@@

<!-- This section presents a new length-weight relationship for lionfish in the central Mexican Caribbean region. A summary of length-weight relationship is provided, and a comparison of predicted-to-observed weight is included. -->

The model adjusted to Eq. \ref{eq:log-alo-trans} estimated the coefficient values at $b = `r coef(model)[[2]]`$ and $c = `r coef(model)[[1]]`$ ($R^2 = `r Rsq`$, `r Fstring`, $p < 0.001$). The allometric factor ($b$) was significantly different from the value of isometric growth of $b = 3$ ($t(107) = `r formatC((coef(model)[[2]]-3)/robust_se[[2]], digits = 2, format = "f")`; p<0.001$), indicating that lionfish present allometric growth. More information on model fit is presented in TableS2. The relationship between TL and TW is presented in Figure \ref{fig:l-w-carib}.

From this study in the central Mexican Caribbean and the `r length(unique(review$Study))-1` peer-reviewed studies that reported length-weight parameters for *P. volitans* `r dim(review)[1]-1` parameters were identified (Table \ref{tab:all_params}, Fig \ref{fig:all_allo}). Two studies \citep{aguilarperera_2016,fogg_2013} reported gender-level and pooled parameters, while the rest presented pooled results. Reviewed studies presented information for organisms obtained at depths between `r min(review$minD, na.rm = T)` and `r max(review$maxD, na.rm = T)` m. Three studies explicitly stated that their organisms were sampled with pole spears \citep{aguilarperera_2016,chin_2016,dahl_2014,sabidoitz_2016}, and five studies mentioned that some of their organisms were obtained with pole spears (or other type of harpoon) but also hand-held nets or fish traps \citep{sandel_2015,barbour_2011,fogg_2013,edwards_2014,sabidoitza_2016,sabidoitz_2016,toledohernndez_2014}, and two studies did not specify how organisms were sampled \citep{deleon_2013,darling_2011}. \citet{fogg_2013} use spine-less weight in the length-weight relationship estimation.

@@

There were significant differences in predicted-to-observed weight ratios estimated for each pair of parameters (F(df = 15; 1728) = 38.26; p < 0.001). From all allometric parameters reviewed, those of \citet{sabidoitz_2016} in Banco Chinchorro, (Caribbean) provided the lowest weight estimates, with a predicted-to-observed weight ratio of `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` (mean $\pm$ SD). On the other hand, \citet{barbour_2011} in the Northern Atlantic yielded the highest weight estimates, with a mean ($\pm$ SD) predicted-to-observed weight ratio of `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")`. Predicted-to-observed weight ratios are presented in Figure \ref{fig:bio_ratio}.

@@

review %>% 
  filter(!Study == "This study") %>% 
  mutate(Length = list(seq(1, 400))) %>% 
  unnest() %>% 
  mutate(Weight = a*(Length^b),
         This = 10^(coefficients(model)[1] + coefficients(model)[2]*log10(Length))) %>%
  group_by(Length) %>% 
  mutate(min = min(Weight),
         max = max(Weight)) %>% 
  ungroup() %>% 
  ggplot(aes(x = Length)) +
  geom_line(aes(y = Weight, color = Study, linetype = Gender, group = Study_id), size = 1) +
  # geom_ribbon(aes(ymin = min, ymax = max), color = "gray", alpha = 0.1) +
  geom_line(aes(y = This), size = 1.2, color = "black", linetype = "dashed") +
  labs(x = "Total length (mm)", y = "Total Weight (gr)") +
  theme_cowplot() +
  theme(text = element_text(family = "Times"),
        legend.justification=c(0,1),
        legend.position=c(0.01, 1),
        legend.background = element_rect(fill = "transparent"),
        legend.box = "horizontal",
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10)) +
  guides(color = guide_legend(order = 1),
         linetype = guide_legend(order = 0, ncol = 2),
         size = "none") +
  scale_color_brewer(palette = "Paired", direction = -1) +
  scale_linetype_manual(values=c("solid", "longdash", "dotted")) +
  scale_size_manual(values = c(1, 1.5))
```

```{r, fig.height = 4.2, fig.width = 5.5, fig.cap = "\\label{fig:bio_ratio}Violin plot showing the distribution of predicted to observed weight ratios for 18 pairs of allometric parameters. Red and blue circles indicate median and mean values, respectively. Like letters indicate values that do not differ significantly (Tukey's HSD; p < 0.05)."}
groups$groups%<>%
  mutate(Study_id = rownames(.)) %>% 
  select(-Biomass_ratio)

@@

A new pair of allometric growth parameters for lionfish in the central Mexican Caribbean are provided, complimenting existing literature for other sites in the Mexican Caribbean \citep{sabidoitza_2016,sabidoitz_2016}. Additionally, the study identifies regional differences in length-weight relationships.

The length-weight coefficients estimated in this study were within the range identified by studies in other regions (Table \ref{tab:all_params}). However, the ones presented here provide lower weight estimates for similar lenmgths. Until about TL = 200 mm, there are no appreciable differences between the parameters for organisms from the Mexican Caribbean and those for little Cayman \citep{edwards_2014} and Jamaica \citep{chin_2016}. Yet, for larger organisms (TL > 270 mm) parameters from Costa Rica \citep{sandel_2015} and Bonaire \citep{deleon_2013} provide similar estimates to those from this study. Conversely, these same studies tend to estimate higher weights --as compared to the ones reported here-- for smaller organisms, likely due to the lack of small organisms in the samples used to estimate their parameters.

There are evident differences in weight-at-length between organisms from the Caribbean and Gulf of Mexico / North-Western Atlantic. Weight estimates with parameters from the Gulf of Mexico and North-Western Atlantic tend to be higher than those from the Caribbean. Similar regional variation has been reported for age-at-length relationships of this species across the invaded region \citep{fogg_2015,edwards_2014}, or when comparing populations from the invasion and native ranges \citep{pusack_2016}. These may be driven by genetic differences or organisms being exposed to distinct environmental conditions. For example, work on mitochondrial DNA has shown two distinct population groups, identified as the "Caribbean group" and "Northern Group" \citep{betancurr_2011}. Alternatively, \citet{fogg_2015} suggest that differences observed in age-at-length "may bemore related to climate rather than other biological and ecological factors". Differences in weight-at-length could also reflect differences in energy input (*i.e.* in some regions, lionfish eat more) or differential usage of this energy (*e.g.* regional differences in predator abundances lead to different usage of energy), or a combination of both. Future research should focus on identifying which occurs here.

The results presented in this paper can have major implications for management. For example, \citet{edwards_2014} simulate lionfish culling using parameters from North Carolina and Little Cayman, and identify that the difference in time required for the population to recover to 90% of its initial biomass after removals cease was of up to four years. Our results show that, for a given length, using one set of parameters or the other can result in a threefold increase in estimated weight. This differences become especially relevant when estimating biomass available for harvest, predicting effects on local ecosystems, or evaluating the effectiveness of removal programs. Future research should try to use, to the extent possible, parameters calculated for their region, or use different parameters to provide upper and lower bounds in their results. At the same time, this highlights the need for more basic research that furthers our understanding of lionfish biology. To better manage the invasion, we must perform research that can describe biologically important information of lionfish throughout its invasion range \citep{johnson_2016}.

@@

I would like to thank thank Nils Van Der Haar and Michael Doodey from Dive Aventuras as well as Guillermo Lotz-Cador who provided help to collect samples.

>>>>

---
title: Invasive lionfish present region-wide variation of allometric growth in the
  Western Atlantic
author: "Juan Carlos Villaseñor-Derbez$^1$, Sean Fitzgerald"
header-includes:
- \usepackage{float}
- \floatplacement{figure}{H}
output:
  pdf_document:
    keep_tex: yes
    number_sections: no
    template: peerj_template.tex
keywords: A, B, C
institution: Bren School of Environmental Sciences and Management, University of California
  Santa Barbara, Santa Barbara, California, USA
abstract: "Lionfish (*Pterois volitans/miles*) are an invasive species in the Western Atlantic and the Caribbean. Improving management of invasive lionfish populations requires accurate estimates of total biomass, which in turn depend on accurate definition of the length-weight relationship. Here, we reviewed published length-weight relationships of lionfish taken from throughout their invasive range and found that lionfish of equal lengths have lower body mass in the Caribbean than in the Atlantic or Gulf of Mexico. Additionally, we report a new pair of length-weight parameters ($a = 3.2056 \\times 10^{-6}; b = 3.235$) for organisms sampled in the central Mexican Caribbean region. The substantial spatial variation  in length-weight parameters highlights the importance of using site-specific information when estimating lionfish biomass using length observations. These findings can have major implications for management in terms of predicting effects on local ecosystems, evaluating the effectiveness of removal programs, or estimating biomass available for harvest."
---

@@

Lionfish (*Pterois volitans/miles* complex) are an invasive species in the western Atlantic and the Caribbean, likely introduced through liberation of aquarium-kept organisms \citep{betancurr_2011}. They are the first invasive marine vertebrates established along the North Atlantic Caribbean coasts \citep{schofield_2009,schofield_2010,sabidoitza_2016} and their presence has been labeled as a major marine invasion because they threaten local biodiversity, spread rapidly, and are difficult to manage \citep{hixon_2016}. Invasive lionfish are primarily studied in coral reef ecosystems, where their impacts are far-reaching. For example, field experiments by \citet{albins_2008} showed that lionfish establishment led to reduced recruitment of native fishes by nearly 80% over a five week period in Florida, and increased lionfish biomass coincided with a 65% reduction in the biomass of native prey fishes along Bahamian coral reefs in just two years \citep{green_2012}. Lionfish have also established invasive populations in other habitats such as estuaries, mangroves , hard-bottomed areas , and mesophotic reefs \citep{jud_2011,barbour_2010,muoz_2011,andradibrown_2017,claydon_2012}.

A substantial amount of research describes lionfish feeding ecology from North Carolina to Costa Rica \citep{muoz_2011,morris_2009,cote_2013,dahl_2014,valdezmoreno_2012,villaseorderbez_2014,hackerott_2017,sandel_2015}. A meta-analysis by  \citet{peake_2018} showed that invasive lionfish prey on at least 167 different species across the tropical and temperate North Atlantic. Their feeding behavior and high consumption rates can reduce recruitment and population sizes of native reef-fish species, and can further endanger reef fish \citep{albins_2008, green_2012,rocha_2015}. (However, see \citet{hackerott_2017} for a case where there was no evidence that lionfish affected the density, richness, or community composition of prey fishes). Major efforts have been made to understand the possible impacts of the invasion by tracking the spread of established lionfish populations through time \citep{schofield_2009,schofield_2010} and by predicting invasion ranges under future climates \citep{grieve_2016}. Trophic impacts of lionfish can be minimized if local lionfish biomass is controled by by culling \citep{ariasgonzalez_2011}. However, the feasability of controling local populations remains a topic of debate \citep{cote_2014}.

Governments and non-profit organizations have sought to reduce lionfish densities through removal programs and incentivizing its consumption \citep{chin_2016}. In some cases, these have shown to significantly reduce --but not quite eliminate-- lionfish abundances at local scales \citep{sandel_2015,chin_2016,deleon_2013}. In addition, culling programs can help stabilize or grow native prey fish populations \citep{cote_2014}. Complete eradication of lionfish through fishing is unlikely because of their rapid recovery rates and ongoing recruitment to shallow-water areas from their persistent populations in mesophotic coral ecosystems \citep{barbour_2011,andradibrown_2017}. However, promoting lionfish consumption might create a level of demand capable of sustaining a stable fishery, which can help control shallow-water populations while providing alternative livelihoods and avoiding further impacts to local reef biota \citep{chin_2016}.

The feasibility of establishing fisheries through lionfish removal programs has been extensively evaluated through field observations and empirical modeling \citep{barbour_2011,morris_2011,deleon_2013,johnston_2015,sandel_2015,chin_2016,usseglio_2017}. One contributing factor to the success of many removal programs is the sedentary nature of adult lionfish \citep{jud_2012}. Culling programs are effective in reducing adult populations largely because lionfish exhibit high levels of site fidelity and rarely leave their home range in most cases \citep{Fishelson_1997,cote_2014,kochzius_2005} -- But see \citet{andradibrown_2017} for cases when deep-water populations maintain shallow-water populations. Fish with this sedentary behavior are likely to exhibit high levels of spatial variation in important life history characterstics such as growth or natural mortality rate \citep{hutchinson_2008,wilson_2012}. The importance of considering spatial heterogeneity is well-documented in terms of assessing and managing sedentary species \citep{gunderson_2008,guan_2013}, and such variation should be accounted for when evaluating the feasibility of establishing lionfish fisheries as well.

Empirical modeling efforts examining the feasibility of establishing fisheries for lionfish involve modeling changes in biomass in response to changes in mortality (*i.e.* culling). A common way to model this is via length-structured population models, where fish lengths are converted to  weight in order to calculate total biomass \citep{cote_2014,barbour_2011,andradibrown_2017}. The length-weight relationship is therefore an essential component of these models, but this relationship can vary across regions as a response to biotic (e.g. local food availability) and abiotic (e.g. water temperature) conditions \citep{johnson_2016}. Literature suggests that site-specific parameters are necessary in order to accurately estimate biomass when length-weight relationships are spatially variable , and this variability becomes increasingly important when estimating the potential effectiveness of (and resources needed for)  lionfish culling programs or when identifying total biomass available for harvest by fishers \citep{barbour_2011,morris_2011,johnston_2015,chin_2016,cote_2014}. In addition to environmentally-driven spatial variation, genetic analysis of invasive lionfish suggest biological differences due to the existence of two genetically distinct subpopulations between the northwest Atlantic and the Caribbean \citep{betancurr_2011}. To date, no studies have examined region-wide differences in length-weight parameters despite the large number of studies reporting this relationship for lionfish.

The objective of this paper is to describe the spatial pattern of length-weight relationships of lionfish in the Caribbean and Western Atlantic and evaluate the implications of these spatial differences. Length-weight relationships for lionfish exist for North Carolina, Northern and Southern Gulf of Mexico, the Southern Mexican Caribbean, Bahamas, Little Cayman, Jamaica, Bonaire, Puerto Rico, and Costa Rica \citep{barbour_2011,fogg_2013,dahl_2014,aguilarperera_2016,sabidoitza_2016,sabidoitz_2016,darling_2011,edwards_2014,chin_2016,deleon_2013,toledohernndez_2014,sandel_2015}. This study also provides the first length-weight relationship for the central Mexican Caribbean.

@@

We reviewed 12 published studies and obtained 17 length-weight relationships for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and Caribbean (n = 10, Table \ref{tab:all_params}, Fig \ref{fig:all_allo}). We collected information on sampling methods, sex differentiation, location, and depth ranges from each study when available, and assumed both genders were included in a study if gender was unspecified. Two studies reported parameters for genders combined an separated \citep{aguilarperera_2016,fogg_2013}, while the rest presented pooled results. Reviewed studies presented information for organisms obtained at depths between `r min(review$minD, na.rm = T)` and `r max(review$maxD, na.rm = T)` m. Three studies explicitly stated that their organisms were sampled with pole spears \citep{aguilarperera_2016,chin_2016,dahl_2014,sabidoitz_2016}, and five studies mentioned that some of their organisms were obtained with pole spears (or other type of harpoon) but also hand-held nets or fish traps \citep{sandel_2015,barbour_2011,fogg_2013,edwards_2014,sabidoitza_2016,sabidoitz_2016,toledohernndez_2014}, and two studies did not specify how organisms were sampled \citep{deleon_2013,darling_2011}. \citet{fogg_2013} use spine-less weight in the length-weight relationship estimation, and thus their parameters likely underestimate whole wieght. Since no spine-less to whole weight conversions were available, these parameters were taken as reported.

We also collected data from 10 sampling sites along the central Mexican Caribbean coast in 2010 (Table S1). Sampling locations included wall and carpet reefs at depths between 5.7 m and 38.1 m. All observed lionfish (n = 109) were collected using hand nets and numbered collection bottles. The use of hand nets prevented any weight loss due to bleeding and allowed better representation of small sizes by eliminating gear selectivity. Organisms were euthanized via pithing and Total Length (TL; mm) and Total Weight (TW; g) were recorded.

@@

@@

\begin{equation}
\label{eq:log-alo-trans}
Y = bX + c
\end{equation}

Where $Y = log_{10}(TW)$, $X = log_{10}(TL)$, and $c = log_{10}(a)$. The coefficients ($c$ and $b$) were estimated with  an Ordinary Least Squares Regression and heteroskedastic-robust standard error correction \citep{zeileis_2004}. The $b$ coefficient was tested against the null hypothesis of isometric growth (*i.e.* $H_0: b = 3$). Coefficients were tested with a two-tailed Student's t, and the significance of the regression was corroborated with an F-test.

Some of the reviewed studies inconsistently defined $a$ as either the ponderal index from Eq. \ref{eq:allometric} or the y-intercept ($c$) from Eq. \ref{eq:log-alo-trans}. Other studies incorrectly reported parameters as mm-to-g conversions when they were in fact cm-to-g conversions. We standardized each study by converting coefficients and report all parameters as TL(mm) to TW (gr) conversions. Locations where allometric studies have been performed are shown in Figure \ref{fig:map} and Table \ref{tab:all_params}.

@@

Combining the length-weight parameters extracted from the literature and the additional pair calculated here, we obtain a total of 18 pairs. We use the central Mexican Caribbean as a case study of how the use of *ex situ* parameters influences the accuracy of weight estimates for lionfish. We estimated TW from the TL observations we collected in the central Mexican Caribbean (n = 109) using each of the 18 parameter pairs and divided predicted weights by known observed weights to obtain a simple measure of over- or underestimation. Difference in mean weight ratios across the different parameter pairs were tested with a one-way analysis of variance (ANOVA). All analyses were performed in R version `r paste(version$major, version$minor, sep = ".")` \citep{rcore_2018} . Raw data and code used in this work are available at dryad.org.

@@

Inserting TL observations for the central Mexican Caribbean into \ref{eq:log-alo} Eq. 2 and converting to \ref{eq:log-alo-trans} resulted in the coefficient values $b = `r b`$ and $c = `r c`$ ($R^2 = `r Rsq`$, `r Fstring`, $p < 0.001$). The length-weight coefficients estimated in this study were within the range identified by studies in other regions (Table \ref{tab:all_params}). The significant difference between this allometric factor ($b$) and the isometric growth factor $b = 3$ ($t(107) = `r formatC((coef(model)[[2]]-3)/robust_se[[2]], digits = 2, format = "f")`; p<0.001$) indicated that lionfish present allometric growth, which is consistent with other studies. Figure \ref{fig:l-w-carib} shows the relationship between TL and TW for this region, and more information on model fit is presented in Table S2.

@@

There were significant differences in our predicted weights for the central Mexican Caribbean when using the different pairs of parameters ($F(df = 15; 1728) = 38.26; p < 0.001$). The lowest weight estimates resulted from using the allometric parameters from Banco Chinchorro in the Caribbean \citet{sabidoitz_2016}, and the highest weight estimates came from the Northern Atlantic\citet{barbour_2011}. The calculated ratio of predicted-to-observed weight  ranged from `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` to `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` (mean $\pm$ SD). Predicted-to-observed weight ratios are presented in Figure \ref{fig:bio_ratio}. Spine-less weight parameters from \citet{fogg_2013} still produced predicted-to-observed weight rations > 1.

@@

review %>% 
  filter(!Study == "This study") %>% 
  mutate(Length = list(seq(1, 400))) %>% 
  unnest() %>% 
  mutate(Weight = a*(Length^b),
         This = 10^(coefficients(model)[1] + coefficients(model)[2]*log10(Length))) %>%
  group_by(Length) %>% 
  mutate(min = min(Weight),
         max = max(Weight)) %>% 
  ungroup() %>% 
  ggplot(aes(x = Length)) +
  geom_line(aes(y = Weight, color = Study, linetype = Gender, group = Study_id), size = 1) +
  geom_line(aes(y = This), size = 1.2, color = "black", linetype = "dashed") +
  labs(x = "Total length (mm)", y = "Total Weight (gr)") +
  theme_cowplot() +
  theme(text = element_text(family = "Times"),
        legend.justification=c(0,1),
        legend.position=c(0.01, 1),
        legend.background = element_rect(fill = "transparent"),
        legend.box = "horizontal",
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10)) +
  guides(color = guide_legend(order = 1),
         linetype = guide_legend(order = 0, ncol = 2),
         size = "none") +
  scale_color_brewer(palette = "Paired", direction = -1) +
  scale_linetype_manual(values=c("solid", "longdash", "dotted")) +
  scale_size_manual(values = c(1, 1.5))
```

```{r, fig.height = 4.2, fig.width = 5.5, fig.cap = "\\label{fig:bio_ratio}Violin plot of predicted to observed weight ratios for 18 pairs of allometric parameters. Red and blue circles indicate median and mean values, respectively. Like letters indicate values that do not differ significantly (Tukey's HSD; p < 0.05)."}
groups$groups%<>%
  mutate(Study_id = rownames(.)) %>% 
  select(-Biomass_ratio)

@@

We  detected substantial differences in weight-at-length between organisms from the Caribbean and Gulf of Mexico / North-Western Atlantic. Weight estimates using parameters from the Gulf of Mexico and North-Western Atlantic were higher on average than those from the Caribbean. The average predicted-to-observed weight ratios from these three regions were insert 1.24 $\pm$ 0.309, 1.76 $\pm$ 0.496, and Caribbean 1.17 $\pm$ 0.398, respectively. These length-weight differences mirror similar findings of regional variability in age-at-length relationships of lionfish across both their invaded and native regions \citep{pusack_2016}. These differences may be driven by genetic variation or by organisms being exposed to distinct environmental conditions. For example, \citet{betancurr_2011} used mitochondrial DNA to demonstrate the existence of two distinct population groups, identified as the “Caribbean group” and “Northern Group”, and \citet{fogg_2015} alternatively suggested that age-at-length differences may be driven by climate. Differences in weight-at-length could also reflect differences in energy input (*i.e.* in some regions, lionfish eat more) or differential usage of this energy (*e.g.* regional differences in predator abundances lead to different usage of energy), or a combination of both. Future research is needed to determine which processes are at work here.

Differences in length-weight relationships have traditionally been highlighted as potential pitfalls to fishery management. For example, \citet{wilson_2012} show that small-scale variations in length-at-age and fishing mortality in other Scorpaeniformes (*Sebastes rastrelliger*) translate to differential landings, effort, and catch per unit effort in the live fish fishery of California, and that these differences must be taken into account in management plans. The lionfish case poses the opposite scenario, where the manager desires to erradicate species. To accurately gauge both the effectiveness of lionfish removal efforts and the resources needed to successfully manage an invasion, we must acknowledge and understand regional biological differences in important variables such as allometric growth parameters.

The results presented here have major implications for management. For example, \citet{edwards_2014} simulated a lionfish culling program under two scenarios, one using length-at-age and length-to-weight parameters from North Carolina and one using parameters from Little Cayman. They showed that using the different parameters caused up to a four-year difference in the time required for the simulated lionfish population to recover to 90% of its initial biomass after removals ceased. Here, we show that using one set of length-weight parameters versus another for a given length can result in up to a threefold TW overestimation. These differences become especially important when allocating resources for removal programs, incentivizing its fishery as alternative livelihoods, or estimating ecosystem impacts. Research efforts focused on invasive lionfish populations need to use parameters calculated for their region to the extent possible, or at least use reasonable sets of different parameters that provide upper and lower bounds in their results. This work additionally highlights the need for more basic research that furthers our understanding of the invasive lionfish.

@@

The authors would like to thank thank Nils Van Der Haar and Michael Doodey from Dive Aventuras as well as Guillermo Lotz-Cador who provided help to collect samples.
